An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-31 Revision/response.tex

To make the paper more readable, we have added a more detailed explanation of the range attribute in Section 4.3. \textbf{(Please see Section 4.3, highlighted in blue in the revised manuscript.)}

\textcolor{blue}{"Due to the limitations of existing range filtering algorithms, our experiment uses only a single attribute, with the dataset sorted in ascending order by value. The WIT-Real and YT-Audio-Real datasets use real attribute values, while the rest use randomly generated ones."}

In response to the large difference in index construction time shown in Figure 12a, we have revised the paper based on your suggestion. We added more detailed and in-depth analysis to provide a clearer explanation. \textbf{(Please see the "Index construction time" part of Section 4.3.1, highlighted in blue in the revised manuscript.)}

	\textcolor{blue}{
	"As shown in Figure 12a, Faiss and Milvus, based on the IVF\_Flat architecture, require less time. This indexing structure enables fast construction through subset sampling and centroid computation, thereby reducing computational cost. Milvus performs particularly well on large-scale datasets, mainly due to its cache-aware optimizations, which significantly reduce CPU cache misses.}

	\textcolor{blue}{In addition, the two HNSW-based methods, ACORN-1 and SeRF, also achieve short index construction time. ACORN-1 closely follows the original HNSW construction process and reduces the number of candidate neighbors in the upper layers to lower computational overhead. SeRF attaches timestamp information to edges during construction, making the total build time comparable to that of a single HNSW graph.
	}

	\textcolor{blue}{We observe that ACORN-\(\gamma\), DSG and WinFilter exhibit relatively long index construction time. Although the first two methods are also based on HNSW, ACORN-\(\gamma\) extends the neighbor list of each node to provide more candidate paths, which increases both computational and memory overhead. DSG, on the other hand, incorporates a rectangle tree structure for index management, making the construction process more complex and costly. 
		WinFilter also incurs high construction cost, as it relies on a tree structure and builds a separate nearest neighbor graph for each node, which leads to redundant computations. However, on large-scale datasets, its preprocessing stage is replaced by the base algorithm Vamana-WST, effectively avoiding significant time overhead."}

### 2025-07-31 Revision/response.tex

	\caption{Different AS under the Same Algorithm}
	\label{fig:attribute_selectivity}
\end{figure*}
\begin{figure}[htbp]
	\centering
	%	\includegraphics[width=0.8\linewidth]{fig/exp_5_2_1_legend.pdf}
	\includegraphics[width=\linewidth]{fig/exp_5_1_1_SingleLabel_1thread.pdf}
	\caption{Different AS under the Same Algorithm}
	\label{fig:attribute_selectivity_2}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%% R2 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{metarevision}
		\textit{Run experiments with multi-modal queries (e.g., combining text and images) or more complex query types involving both attribute and range filters.}
	\end{metarevision}
}

\noindent
\textbf{Response:} We believe this is an excellent suggestion. To provide a more comprehensive evaluation of the robustness of each algorithm, we introduce new experiments on two challenging multi-modal datasets.

First, we have incorporated a standard text-to-image dataset, which is widely adopted by existing multi-modal ANN algorithms. In this dataset, the base vectors are images while the query set consists of texts, creating a modality gap that demands stronger algorithm robustness to achieve good performance. Second, to further test performance on complex queries, we have constructed a more challenging dataset with a base set of 500,000 text vectors and 500,000 image vectors, and a query set containing 5,000 text and 5,000 image vectors. On this latter dataset, we perform both attribute filtering and range queries using combined text and image inputs to analyze the algorithm's robustness under more complex, mixed-modality conditions. The experimental results are shown in Figure~\ref{fig:attribute_multimodel} and Figure~\ref{fig:range_multimodel}.

As for complex query types that involve both attribute filtering and range queries simultaneously, to the best of our knowledge, existing algorithms—except for general-purpose systems such as Milvus and Faiss—do not support this functionality. For example, ACORN needs to build separate indexes for each scenario and cannot handle attribute filtering and range queries simultaneously. Therefore, we do not consider this type of complex query in our experiments but instead study the two query types separately. \textbf{(Please see Section 4.1.1, Section 4.2.3 and Secntion 4.3.3, highlighted in blue in the revised manuscript.)  
	%Please refer to the blue highlighted parts in  Section 4.1.1 paragraph 4 and  Section 4.2.4 and Section 4.3.4 under Multi-modal Queries.)
}

Section 4.1.1:\textcolor{blue}{
"Additionally, we use the Out-of-Distribution (OOD) dataset Text2Image and the large-scale dataset Deep100M to evaluate the generalization capability of the algorithms. The Text2Image dataset includes two versions. The original version contains 1 million image embeddings as the base dataset and uses 10 thousand text embeddings as the query set. The Text2Image-Mix version is composed of 0.5 million image embeddings and 0.5 million text embeddings as the base dataset, with a query set of 5 thousand image embeddings and 5 thousand text embeddings."
}

@@

Section 4.2.3: \textcolor{blue}{
	"We conduct experiments on two multimodal datasets Text2image-Mix and Text2image. On Text2imageMix dataset, as shown in Figure 9, UNG and NHQ consistently perform best, followed by CAPS. Although the dataset contains two modalities, due to the significant differences between the modalities, they are actually two separate regions in the vector space. The performance is similar to that of the single-modal dataset in Figure 4. This is because mixed multi-modal queries are actually the same as conducting experiments on two datasets of the same modality. This is not actually a difficult query. In contrast, the Text2image dataset shown in Figure 9 is different. The query set and the base dataset belong to two completely different modalities. Queries are far from the dataset points, and the ground truths are more dispersed. To find the groundtruths, algorithms typically need to examine more points, which leads to degraded performance." 
}

Section 4.3.3:\textcolor{blue}{
		"Similar to attribute filtering, on Text2Image-Mix dataset, the performance of multi-modal range filtering queries shown in Figure 14 also aligns with the single-modality results in Figure 13. These algorithms on the Text2Image dataset exhibit a significant performance degradation. This is because most of these range query algorithms are graph-based, and such algorithms require more iterations to find true neighbors in OOD datasets."
}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.95\linewidth]{fig/attribute_legend.pdf}
	\includegraphics[width=\linewidth]{fig/attribute_multimodel_2.pdf}
	\caption{Multi-modal Datasets (AF)}
	\label{fig:attribute_multimodel}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.95\linewidth]{fig/range_legend.pdf}
	\includegraphics[width=\linewidth]{fig/range_multimodel_2.pdf}
	\caption{Multi-modal Datasets (RF)}
	\label{fig:range_multimodel}
\end{figure}

### 2025-07-31 Revision/response.tex

As for complex query types that involve both attribute filtering and range queries simultaneously, to the best of our knowledge, existing algorithms—except for general-purpose systems such as Milvus and Faiss—do not support this functionality. For example, ACORN needs to build separate indexes for each scenario and cannot handle attribute filtering and range queries simultaneously. Therefore, we do not consider this type of complex query in our experiments but instead study the two query types separately.

Section 4.1.1: \textcolor{blue}{
"Additionally, we use the Out-of-Distribution (OOD) dataset Text2Image and the large-scale dataset Deep100M to evaluate the generalization capability of the algorithms. The Text2Image dataset includes two versions. The original version contains 1 million image embeddings as the base dataset and uses 10 thousand text embeddings as the query set. The Text2Image-Mix version is composed of 0.5 million image embeddings and 0.5 million text embeddings as the base dataset, with a query set of 5 thousand image embeddings and 5 thousand text embeddings."
}

Section 4.2.3: \textcolor{blue}{
	"We conduct experiments on two multimodal datasets Text2image-Mix and Text2image. On Text2imageMix dataset, as shown in Figure 9, UNG and NHQ consistently perform best, followed by CAPS. Although the dataset contains two modalities, due to the significant differences between the modalities, they are actually two separate regions in the vector space. The performance is similar to that of the single-modal dataset in Figure 4. This is because mixed multi-modal queries are actually the same as conducting experiments on two datasets of the same modality. This is not actually a difficult query. In contrast, the Text2image dataset shown in Figure 9 is different. The query set and the base dataset belong to two completely different modalities. Queries are far from the dataset points, and the ground truths are more dispersed. To find the ground truths, algorithms typically need to examine more points, which leads to degraded performance." 
}

@@

To make the paper more readable, we have added a more detailed explanation of the range attribute in Section 4.3. \textbf{(Please see Section 4.3, highlighted in blue in the revised manuscript.)}

\textcolor{blue}{"Due to the limitations of existing range filtering algorithms, our experiment uses only a single attribute, with the dataset sorted in ascending order by value. The WIT-Real and YT-Audio-Real datasets use real attribute values, while the rest use randomly generated ones."}

In response to the large difference in index construction time shown in Figure 12a, we have revised the paper based on your suggestion. We added more detailed and in-depth analysis to provide a clearer explanation. \textbf{(Please see the "Index construction time" part of Section 4.3.1, highlighted in blue in the revised manuscript.)}

\textcolor{blue}{
	"ACORN-1 also achieves highly efficient index construction, especially on large-scale datasets. This efficiency stems from its construction process, which closely resembles that of the original HNSW, and from reducing the number of candidate neighbors in the upper layers, thereby lowering computational cost. SeRF, another HNSW-based method, also exhibits very short construction time. During construction, it attaches timestamp information to edges, keeping the total time comparable to building a single HNSW graph, which enables efficient construction even on large datasets.}

\textcolor{blue}{
	In contrast, WinFilter relies on a tree-based structure, where a separate nearest neighbor graph must be built for each node. This design introduces redundant computation and high construction cost. However, on large-scale datasets, the preprocessing phase of WinFilter is replaced in the experiments by the base algorithm Vamana-WST, thus avoiding significant time overhead.}

\textcolor{blue}{
	ACORN-\(\gamma\)  extends the neighbor list of each node to provide more candidate paths, which increases both computational and memory overhead. Although it prunes distant neighbors to save storage, the additional construction logic still incurs substantial time cost.}

	\textcolor{blue}{
	DSG is an optimization over SeRF, incorporating a rectangular tree structure for index management. This makes the construction process more complex and costly. In addition, on Deep100M, DSG uses single-threaded construction, resulting in the longest build time."}

### 2025-07-31 VLDB2026/main.tex
gt

		\item \textbf{Recall@k}: The proportion of overlap between the returned approximate $k$ nearest neighbors and the ground-truth $k$ nearest neighbors.
		\item \textbf{QPS (Queries Per Second)}: The number of queries processed per second.
		\item \textbf{Index Construction Time}: The total time required to transform the raw dataset into a queryable index structure.
		\item \textbf{Index Size}: The storage size of the persisted index on disk.
		\item \textbf{Peak Memory Usage}: The maximum memory consumption observed during index construction \textcolor{cyan}{and search}.
	\end{enumerate}

@@

		In contrast, the Text2image dataset shown in Figure~\ref{fig:attribute-multimodal} is different. The query set and the base dataset belong to two completely different modalities. Queries are far from the dataset points, and the ground truths are more dispersed. To find the groundtruths, algorithms typically need to examine more points, which leads to degraded performance.

>>>>

		\item \textbf{Recall@k}: The proportion of overlap between the returned approximate $k$ nearest neighbors and the ground truth $k$ nearest neighbors.
		\item \textbf{QPS (Queries Per Second)}: The number of queries processed per second.
		\item \textbf{Index Construction Time}: The total time required to transform the raw dataset into a queryable index structure.
		\item \textbf{Index Size}: The storage size of the persisted index on disk.
		\item \textbf{Peak Memory Usage}: The maximum memory consumption observed during index construction \textcolor{cyan}{and search}.
	\end{enumerate}

@@

		In contrast, the Text2image dataset shown in Figure~\ref{fig:attribute-multimodal} is different. The query set and the base dataset belong to two completely different modalities. Queries are far from the dataset points, and the ground truths are more dispersed. To find the ground truths, algorithms typically need to examine more points, which leads to degraded performance.
